### 2023-11-09 test.qmd
add new files

---
title: "practice"
format: html
editor: visual
crossref:
  custom: 
    - kind: float # crossreferenceable elements with captions are `float`
      prefix: Supplementary Figure # used as the prefix for reference in output ("In Figure 1, ..")
      name: Supplementary Figure # used as the prefix for captions ("Figure 1: ..")
      ref-type: sfig # used as the prefix for the reference identifier ("In @fig-1, ...")
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed) see @fig-test.

```{r}
#| label: fig-test
library(ggplot2)
ggplot(mtcars, aes(hp, mpg)) + geom_point()
```

random text see @sfig-test2.

```{r}
#| label: sfig-test2
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) + geom_point()
```