### 2023-11-15 compare.Rmd
comparing heterosexual models

---
title: "compare"
author: "C. B. Smith"
date: '2023-11-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(psych)
library(xtable)
library(knitr)
library(kableExtra)
library(weights)
library(jtools)
library(reshape2)
library(gtsummary)
library(corrplot)
library(patchwork)
source('modelsregressiontables.R')
```

Models of lifetime sex partners with original sample, restricted sample, and heterosexual control
```{r partners}
lifetime_compare
```

Models of lifetime sex partners with original sample, restricted sample
```{r past year}
pastyear_compare
```

Models of lifetime sex partners with original sample, restricted sample
```{r age first sex}
agefirst_compare
```

Models of lifetime sex partners with original sample, restricted sample, and heterosexual control
```{r partnered}
partnered_compare
```

```{r married}

Married_compare2
```